---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Counternull

<!-- badges: start -->
<!-- badges: end -->

Counternull package allows users to resample data for
permutation testing. Users may generate null distributions from their own
datasets and calculate tests statistics and P-values. This package allows
users to identify counternull values in a dataset and plot counternull
distributions. The ability for users to obtain both P-values and
counternull values from a dataset helps prevent against common
misinterpretations in hypothesis testing.

## Installation

You can install the released version of Counternull from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("Counternull")
```

## Usage

Examples of functions that can be used in Counternull Package:

```{r example}
library(Counternull)
## prints test statistic and P-value from dataset
## plots null distribution of given dataset
## dashed black line: test statistic
create_null_distribution(sample_district_1DS,0,rand_matrix_1DS,
sample_district_1DS$charge_prosecuted_1000_rate_post,
sample_district_1DS$charge_prosecuted_1000_rate_pre,128)

## prints counternull values of given data set
## plots null and counternull distribution from dataset
## dashed black line: test statistic,
## dashed red line: counternull value,
## dashed gray line: 0
find_counternull_values(.375,sample_district_1DS,0,rand_matrix_1DS,-8000,0,
find_test_stat(sample_district_1DS,
sample_district_1DS$charge_prosecuted_1000_rate_post,
sample_district_1DS$charge_prosecuted_1000_rate_pre),
sample_district_1DS$charge_prosecuted_1000_rate_post,
sample_district_1DS$charge_prosecuted_1000_rate_pre,128,7)

```

## License
[MIT](https://choosealicense.com/licenses/mit/)

