---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Counternull

<!-- badges: start -->
<!-- badges: end -->

Counternull package allows users to conduct Randomization-Based Inference for 
customized experiments. The package may be used to compute Fisher-Exact 
P-Values alongside null randomization distributions. Additionally, 
users can retrieve counternull sets, generate counternull distributions,
compute Fisher Intervals, and Fisher-Adjusted P-Values. The package may be used
on data of any size and distribution including usage with custom made test 
statistics.


## Installation

You can install the released version of Counternull from 
[CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("Counternull")
```

## Usage

Examples of functions that can be used in Counternull Package:

```{r example}
library(Counternull)
y = sample_data$turn_angle
w = sample_data$w
n_r = create_null_rand(y, w, sample_matrix, test_stat = c("t"))
summary(n_r)
plot(n_r)

n_r = create_null_rand(sample_data$turn_angle, sample_data$w,
sample_matrix, test_stat = c("diffmeans"))
c = find_counternull_values(n_r)
summary(c)
plot(c)
```

## License
[MIT](https://choosealicense.com/licenses/mit/)

